```{r setup}
#| message: false
#| include: false
library(tidyverse)
```

# Analyses

This chapter details all analyses. It could be further split in several chapters later, for the moment this is just a dummy example.

### Dummy example

This example shows the building of intermediary files that are stored in the `outputs/` folder to be shared upon publication. Here the intermediary file is simply species abundance in plot 6 from Paracou in 2016 using dummy data.

```{r testabund}
#| message: false
read_tsv("data/dummy_data.tsv") %>% 
  group_by(Plot, Family, Genus, Species) %>% 
  summarise(abundance = n()) %>% 
  write_tsv("outputs/species_abundance.tsv")
```

And then intermediary files can be used to build analyses, such as the rank abundance curves figure shown here.

```{r testrankabund}
#| message: false
#| fig-cap: "Paracou Plot 6 rank abundance curve in 2016."
read_tsv("outputs/species_abundance.tsv") %>% 
  arrange(desc(abundance)) %>% 
  mutate(rank = 1:n()) %>% 
  ggplot(aes(rank, abundance)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  scale_y_log10()
```
