```{r setup}
#| message: false
#| include: false
library(tidyverse)
```

# Logging {.unnumbered}

All site and plot logging data needed for the analyses. They rely on Mithila google sheet "bioforest-plot-metadata".

```{r logprep}
#| eval: false
damages <- read_tsv("outputs/gathered.tsv") %>% 
  select(site, plot, ba_pre, ba_post, delta_ba, delta_ba_pct) %>% 
  unique()
logging <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1fq2owxMBLBwwibcdw2uQQFxnIhsMbaH4Qcj_xUwVvSQ/edit?gid=0#gid=0")
damages %>% 
  left_join(logging %>% 
              rename_all(tolower)) %>% 
  write_tsv("outputs/logging.tsv")
```

## Treatment

Treatment and resulting damages were coherent in Paracou, but reduced impact logging (RIL) did not seems much effective in Missiones.

```{r treat}
#| message: false
#| warning: false
#| fig-cap: "Treatment effect on disturbance intensity."
read_tsv("outputs/logging.tsv") %>% 
  select(site, plot, treatment, delta_ba) %>% 
  mutate(treatment = gsub("_", " ", treatment)) %>% 
  ggplot(aes(treatment, delta_ba, fill = site)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() +
  theme(axis.title.y = element_blank(), legend.position = "bottom") +
  scale_color_discrete("") +
  ylab(expression(Delta~"Basal area of tree " >=10 ~"cm dbh ["~m^{2}~ha^{-~1}~"]"))
```

## Harvested

There seemed to be inconstancies between harvested basal area, number of trees and volume with resulting damages in maximal lost basal area in the five years.

```{r harv}
#| message: false
#| warning: false
#| fig-cap: "Harvest effect on disturbance intensity."
read_tsv("outputs/logging.tsv") %>% 
  select(site, plot, treatment, delta_ba, delta_ba_pct, 
         ba_harvested, nb_trees_harvested, vol_harvested) %>% 
  gather(log_var, log_val, -site, -plot, -treatment, -delta_ba, -delta_ba_pct) %>% 
  na.omit() %>% 
  mutate(log_var = recode(log_var, 
                          "ba_harvested" = "harvested basal area",
                          "nb_trees_harvested" = "number of trees harvested",
                          "vol_harvested" = "harvested volume")) %>% 
  ggplot(aes(log_val, delta_ba, col = site)) +
  geom_point() +
  theme_bw() +
  coord_flip() + 
  geom_smooth(method = "lm", se = FALSE, col = "black") +
  facet_wrap(~ log_var, scales = "free_y", nrow = 3) +
  theme(legend.position = "bottom") + 
  xlab("") +
  ylab(expression(Delta~"Basal area of tree " >=10 ~"cm dbh ["~m^{2}~ha^{-~1}~"]"))
```

## Other disturbances

Wherever available, we may explore other disturbance such liana infestation index, invasive species, fires, storms (known in Missiones), etc. 
