```{r setup}
#| message: false
#| include: false
library(tidyverse)
```

# Climate {.unnumbered}

We developed a [snakemake](https://snakemake.readthedocs.io/en/stable/) and [conda](https://anaconda.org/anaconda/conda) python workflow for automatic climate data gathering from global databases relying here on [TerraClimate](https://www.climatologylab.org/terraclimate.html). The code is available here: <https://github.com/Bioforest-project/LoggingDiversity/tree/main/data> . Figures below show the resulting variables per month for Paracou and Missiones and a detailed figure for Paracou. This could be compared to site data later on for validation.

```{r prepclim}
#| eval: false
list.files("data/climate/", full.names = TRUE) %>% 
  read_tsv() %>% 
  group_by(site) %>% 
  mutate(month = 1:12) %>% 
  mutate(aet = aet*.1, def = def*.1, pdsi = pdsi*.01, pet = pet*.1,
         tmmn = tmmn*.1, tmmx = tmmx*.1, vpd = vpd*.01) %>% 
  write_tsv("outputs/climate.tsv") 
```

```{r allplot}
#| message: false
#| warning: false
#| fig-cap: "Paracou and Missiones climate."
read_tsv("outputs/climate.tsv") %>% 
  gather(variable, value, -site, -month) %>% 
  mutate(var_long = recode(variable,
                           "aet" = "Actual evapotranspiration",
                           "def" = "Climate water deficit",
                           "pdsi" = "Palmer Drought Severity Index",
                           "pet" = "Potential evapotranspiration",
                           "pr" = "Precipitation",
                           "tmmn" = "Minimum temperature",
                           "tmmx" = "Maximum temperature",
                           "vpd" = "Vapor pressure deficit")) %>% 
  ggplot(aes(month, value, col = site)) +
  geom_line() +
  facet_wrap(~ var_long, scales = "free_y") +
  xlab("") + ylab("") +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r betterplot}
#| message: false
#| warning: false
#| fig-cap: "Paracou climate."
read_tsv("outputs/climate.tsv") %>% 
  ggplot(aes(x = month)) +
  geom_col(aes(y = pr/20), fill = "grey", col = NA) +
  geom_line(aes(y = tmmn, col = "maximum")) +
  geom_line(aes(y = tmmx, col = "minimum")) +
  theme_bw() +
  xlab("") + 
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_y_continuous(
    name = "Temperature (°C)",
    sec.axis = sec_axis(trans = ~.*20, name = "Precipitation (mm)")
  ) +
  theme(
    axis.title.y = element_text(color = "black"),
    axis.title.y.right = element_text(color = "grey")
  ) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  facet_wrap(~ site)
```
