```{r setup}
#| message: false
#| include: false
library(tidyverse)
```

# Soil {.unnumbered}

We developed a [snakemake](https://snakemake.readthedocs.io/en/stable/) and [conda](https://anaconda.org/anaconda/conda) python workflow for automatic climate data gathering from global databases relying here on [SoilGrids](https://soilgrids.org/). The code is available here: <https://github.com/Bioforest-project/LoggingDiversity/tree/main/data> . Figures below show the resulting variables per depth for Paracou and Missiones and a detailed figureof textures. This could be compared to site data later on for validation, knowing SoilGrids have know issue for instance in French Guiana.

```{r prepclim}
#| eval: false
list.files("data/soil/", full.names = TRUE) %>% 
  read_tsv() %>% 
  gather(variable, value, -site) %>% 
  separate(variable, c("variable", "depth"), "_") %>% 
  separate(depth, c("mindepth", "depth"), "-") %>% 
  mutate(depth = as.numeric(gsub("cm", "", depth))) %>% 
  select(-mindepth) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  mutate(
    bdod = bdod/100,
    cec = cec/10,
    cfvo = cfvo/10,
    clay = clay/10,
    sand = sand/10,
    silt = silt/10,
    nitrogen = nitrogen/100,
    phh2o = phh2o/10,
    soc = soc/10,
    ocd = ocd/10,
    ocs = ocs/10
  ) %>% 
  write_tsv("outputs/soil.tsv") 
```

```{r allplot}
#| message: false
#| warning: false
#| fig-cap: "Paracou and Missiones soil."
read_tsv("outputs/soil.tsv") %>% 
  gather(variable, value, -site, -depth) %>% 
  ggplot(aes(site, value, fill = depth)) +
  geom_col(aes(group = as.factor(depth)), position = "dodge") +
  facet_wrap(~ variable, scales = "free_y") +
  theme_bw() +
  theme(legend.position = "bottom", axis.title = element_blank()) +
  scale_fill_viridis_c()
```

```{r betterplot}
#| message: false
#| warning: false
#| fig-cap: "Paracou and Missiones soil texture."
read_tsv("outputs/soil.tsv") %>% 
  ggplot(aes(x = sand, y = clay, z = silt, col = site)) +
  ggtern::coord_tern(L = "x", T = "y", R = "z") +
  geom_point() +
  theme_bw() +
  labs(
    yarrow = "Clay (%)",
    zarrow = "Silt (%)",
    xarrow = "Sand (%)"
  ) +
  ggtern::theme_showarrows() +
  ggtern::theme_hidetitles()+
  ggtern::theme_clockwise()
```
