```{r setup}
#| message: false
#| include: false
library(tidyverse)
library(cmdstanr)
library(bayesplot)
```

# Model {.unnumbered}

The model is the same as the modelling framework, but we want to establish a connection between the trajectory of forest structure, as represented by basal area, and forest diversity, as represented by genus diversity.

## Equilibirum

Equilibrium model using control and pre-logging inventories to inform $\theta$ prior of the recovery trajectory:

$$
\begin{align}
BA \sim \mathcal N (\mu_{BA,site}+\delta_{BA,plot}, \sigma_{BA})\\
\delta_{BA,plot} \sim \mathcal N(0,\sigma_{BA,plot}) \\
y \sim \mathcal N (\mu_{site}+\delta_{plot}, \sigma)\\
\delta_{plot} \sim \mathcal N(0,\sigma_{plot}) 
\end{align}
$$

## Recovery

Recovery trajectory model:

$$
\begin{align}
log(BA) \sim \mathcal N (log(\mu_{BA}), \sigma_{BA})\\
\mu_{BA} = \theta_{BA,site} \times (\phi_{BA,plot} + LTP_{BA} \times (1 - \phi_{plot}) + STP_{BA}) \\
LTP_{BA} = 1 - e^{-\lambda_{BA,plot} \times t} \\
STP_{BA} = \delta_{BA,plot} \times (\frac{t}{\tau_{BA,site}} \times e^{1-\frac{t}{\tau_{BA,site}}})^2 \\
log(y) \sim \mathcal N (log(\mu), \sigma)\\
\mu = \theta_{site} \times (\phi_{plot} + LTP \times (1 - \phi_{plot}) + STP) \\
LTP = 1 - e^{-\lambda_{plot} \times t} \\
STP = \delta_{plot} \times (\frac{t}{\tau_{site}} \times e^{1-\frac{t}{\tau_{site}}})^2 \\
\delta_{plot} = f(\phi_{BA,plot})
\end{align}
$$

With the big novelty being:

$$
\delta_{plot} = f(\phi_{BA,plot})
$$

With the function linking the two to be defined.

```{r recovery}
#| warning: false
#| message: false
#| fig-cap: "STP proj with pars."
ltp <- function(time, dist, thetainf, lambda) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  theta0 + (thetainf - theta0) * (ltp)
}
stp <- function(time, dist, thetainf, lambda, tau, delta) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  stp <- delta * (time / tau * exp(1 - time / tau))^2
  theta0 + ltp * (thetainf - theta0) + stp * thetainf
}
data.frame(time = seq(0.1, 50, length.out = 100)) %>%
  mutate(y_ba = stp(time,
    dist = 0.8, thetainf = 30, lambda = 0.05,
    tau = 5, delta = 0
  )) %>% 
  mutate(y_stp = stp(time,
    dist = 0.5, thetainf = 20, lambda = 0.05,
    tau = 5, delta = 0.5
  )) %>%
  mutate(y_ltp = ltp(time, dist = 0.5, thetainf = 20, lambda = 0.05)) %>%
  bind_rows(data.frame(time = 0, y_stp = 20, y_ltp = 20, y_ba = 30)) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = y_ba), col = "#fc8a8a") +
  geom_line(aes(y = y_ltp), col = "grey") +
  geom_line(aes(y = y_stp)) +
  theme_bw() +
  xlab("") + 
  scale_y_continuous(
    name = "Forest diveristy",
    sec.axis = sec_axis(~.*1, name = "Forest structure")
  ) +
  theme(
    axis.title.y.right = element_text(color = "#fc8a8a")
  ) +
  annotate("text",
    x = -4, y = 28, label = "phi[plot]",
    col = "#fc8a8a", parse = TRUE, size = 5
  ) +
    geom_segment(aes(x = -2, y = 25, xend = -2, yend = 30),
    arrow = arrow(length = unit(0.2, "cm")), col = "#fc8a8a"
  ) +
  geom_segment(aes(x = -2, y = 30, xend = -2, yend = 25),
    arrow = arrow(length = unit(0.2, "cm")), col = "#fc8a8a"
  ) +
  annotate("text",
    x = 4, y = 17.5, label = "delta[plot]",
    col = "#a7d6ac", parse = TRUE, size = 5
  ) +
  geom_segment(aes(x = 6, y = 13, xend = 6, yend = 22),
    arrow = arrow(length = unit(0.2, "cm")), col = "#a7d6ac"
  ) +
  geom_segment(aes(x = 6, y = 22, xend = 6, yend = 13),
    arrow = arrow(length = unit(0.2, "cm")), col = "#a7d6ac"
  ) +
  geom_segment(aes(x = -1.5, y = 28, xend = 2, yend = 19),
    arrow = arrow(length = unit(0.2, "cm")), col = "black"
  )
```

## Fit test

```{r}
#| eval: false
recovery <- cmdstan_model("models/recovery.stan")
data <- read_tsv("../modelling/data/derived_data/data_rec.tsv", col_types = cols()) %>%
    select(-y) %>% 
    filter(variable %in% c("ba", "diversity_q1_gen")) %>%
    pivot_wider(names_from = variable, values_from = value) %>% 
    na.omit() %>% 
    mutate(sitenum = as.numeric(as.factor(site))) %>% 
    mutate(plotnum = as.numeric(as.factor(paste(site, plot)))) %>% 
    arrange(sitenum, plotnum, rel_year)
ind <- data %>%
    select(site, plot, sitenum, plotnum) %>%
    unique()
ba_priors <- read_tsv(file.path("..", "modelling",
                                "chains", paste0("equ_", "ba"), "mu.tsv")) %>% 
    select(-sitenum) %>% 
    left_join(select(ind, site, sitenum) %>% unique()) %>% 
    na.omit()
div_priors <- read_tsv(file.path("..", "modelling",
                                 "chains", paste0("equ_", "diversity_q1_gen"),
                                 "mu.tsv")) %>% 
    select(-sitenum) %>% 
    left_join(select(ind, site, sitenum) %>% unique()) %>% 
    na.omit()
mdata <- list(
    n = nrow(data),
    s = max(data$sitenum),
    p = max(data$plotnum),
    ba = data$ba,
    y = data$diversity_q1_gen,
    t = data$rel_year - 3,
    site = data$sitenum,
    plot = data$plotnum,
    site_plot = ind$sitenum,
    mu_ba_theta_s = ba_priors$median,
    sigma_ba_theta_s = ba_priors$sd,
    mu_theta_s = div_priors$median,
    sigma_theta_s = div_priors$sd
)
fit <- recovery$sample(
    mdata,
    chains = 4,
    parallel_chains = 4,
    save_warmup = FALSE,
    refresh = 100
)
fit$save_output_files("chains/test/")
```

## Test result

```{r}
#| message: false
#| warning: false
fit <- as_cmdstan_fit(list.files("chains/test/", full.names = TRUE))
fit$draws(c("gamma_s")) %>% 
  mcmc_dens_chains()
```
