```{r set}
#| message: false
#| include: false
library(tidyverse)
library(scales)
```

# Joint Genus {.unnumbered}

The model is the same as the modelling framework, but we want to establish a connection between the trajectory of forest structure, as represented by basal area, and forest diversity, as represented by genus diversity.

## Equilibirum

Equilibrium model using control and pre-logging inventories to inform $\theta$ prior of the recovery trajectory:

$$
\begin{align}
BA \sim \mathcal N (\mu_{BA,site}+\delta_{BA,plot}, \sigma_{BA})\\
\delta_{BA,plot} \sim \mathcal N(0,\sigma_{BA,plot}) \\
y \sim \mathcal N (\mu_{site}+\delta_{plot}, \sigma)\\
\delta_{plot} \sim \mathcal N(0,\sigma_{plot}) 
\end{align}
$$

## Recovery

Recovery trajectory model:

$$
\begin{align}
log(BA) \sim \mathcal N (log(\mu_{BA}), \sigma_{BA})\\
\mu_{BA} = \theta_{BA,site} \times (\phi_{BA,plot} + LTP_{BA} \times (1 - \phi_{plot}) + STP_{BA}) \\
LTP_{BA} = 1 - e^{-\lambda_{BA,plot} \times t} \\
STP_{BA} = \delta_{BA,plot} \times (\frac{t}{\tau_{BA,site}} \times e^{1-\frac{t}{\tau_{BA,site}}})^2 \\
log(y) \sim \mathcal N (log(\mu), \sigma)\\
\mu = \theta_{site} \times (\phi_{plot} + LTP \times (1 - \phi_{plot}) + STP) \\
LTP = 1 - e^{-\lambda_{plot} \times t} \\
STP = \delta_{plot} \times (\frac{t}{\tau_{site}} \times e^{1-\frac{t}{\tau_{site}}})^2 \\
\delta_{plot} \sim \mathcal N ( \delta0_{site}+\phi_{BA,plot} \times \gamma_{site}, \sigma_{\delta})
\end{align}
$$

With the big novelty being:

$$
\delta_{plot} \sim \mathcal N ( \delta0_{site}+\phi_{BA,plot} \times \gamma_{site}, \sigma_{\delta})
$$

With the function linking the two to be defined.

```{r recovery}
#| warning: false
#| message: false
#| fig-cap: "STP proj with pars."
ltp <- function(time, dist, thetainf, lambda) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  theta0 + (thetainf - theta0) * (ltp)
}
stp <- function(time, dist, thetainf, lambda, tau, delta) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  stp <- delta * (time / tau * exp(1 - time / tau))^2
  theta0 + ltp * (thetainf - theta0) + stp * thetainf
}
g <- data.frame(time = seq(0.1, 30, length.out = 100)) %>%
  mutate(y_ba = stp(time,
    dist = 0.8, thetainf = 30, lambda = 0.05,
    tau = 5, delta = 0
  )) %>%
  mutate(y_stp = stp(time,
    dist = 0.5, thetainf = 20, lambda = 0.05,
    tau = 5, delta = 0.5
  )) %>%
  mutate(y_ltp = ltp(time, dist = 0.5, thetainf = 20, lambda = 0.05)) %>%
  bind_rows(data.frame(time = 0, y_stp = 20, y_ltp = 20, y_ba = 30)) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = y_ba), col = "#fc8a8a") +
  geom_line(aes(y = y_ltp), col = "grey") +
  geom_line(aes(y = y_stp)) +
  theme_bw() +
  xlab("") +
  scale_y_continuous(
    name = "Forest diveristy (e.g. Genus)",
    sec.axis = sec_axis(~ . * 1, name = "Forest structure (e.g. Basal area)")
  ) +
  theme(
    axis.title.y.right = element_text(color = "#fc8a8a")
  ) +
  annotate("text",
    x = -4, y = 28, label = "phi[~plot]",
    col = "#fc8a8a", parse = TRUE, size = 5
  ) +
  geom_segment(aes(x = -2, y = 25, xend = -2, yend = 30),
    arrow = arrow(length = unit(0.2, "cm")), col = "#fc8a8a"
  ) +
  geom_segment(aes(x = -2, y = 30, xend = -2, yend = 25),
    arrow = arrow(length = unit(0.2, "cm")), col = "#fc8a8a"
  ) +
  annotate("text",
    x = 4, y = 17.5, label = "delta[~plot]",
    col = "#a7d6ac", parse = TRUE, size = 5
  ) +
  geom_segment(aes(x = 6, y = 13, xend = 6, yend = 22),
    arrow = arrow(length = unit(0.2, "cm")), col = "#a7d6ac"
  ) +
  geom_segment(aes(x = 6, y = 22, xend = 6, yend = 13),
    arrow = arrow(length = unit(0.2, "cm")), col = "#a7d6ac"
  ) +
  geom_segment(aes(x = -1.5, y = 28, xend = 2, yend = 19),
    arrow = arrow(length = unit(0.2, "cm")), col = "blue"
  ) +
  annotate("text",
    x = 3, y = 23, label = "gamma[~site]",
    col = "blue", parse = TRUE, size = 5
  ) +
  xlab("Years post-diturbance")
ggsave("figures/fig2.png", width = 6, height = 4, g, dpi = 300, bg = "white")
g
```

## Fit

```{r fitres}
#| message: false
#| warning: false
data <- read_tsv("../modelling/data/derived_data/data_rec.tsv",
  col_types = cols()
) %>%
  select(-y) %>%
  filter(variable %in% c("ba", "diversity_q1_gen")) %>%
  select(-sitenum, -plotnum) %>%
  pivot_wider(names_from = variable, values_from = value) %>%
  na.omit() %>%
  filter(!(site %in% c(
    "Manare", "Montagne Tortue", "Nelliyampathy",
    "Uppangala", "BAFOG", "Sao Nicolau",
    "Kabo", "Mil", "Corinto",
    "Peixoto", "Iwokrama", "Antimary", "Peteco"
  ))) %>%
  mutate(sitenum = as.numeric(as.factor(site))) %>%
  mutate(plotnum = as.numeric(as.factor(paste(site, plot)))) %>%
  arrange(sitenum, plotnum, rel_year) %>%
  gather(
    variable, y, -plot, -year, -site, -treatment, -harvest_year,
    -longitude, -harvest_year_min, -rel_year, -sitenum, -plotnum
  )
trajectories <- read_tsv("chains/joint-diversity_q1_gen-trajectories.tsv")
parameters <- read_tsv("chains/joint-diversity_q1_gen-parameters.tsv")
parameters
```

## Trajectories

```{r traj_ba}
#| message: false
#| warning: false
#| fig-cap: "Basal area jointly inferred trajectories for all sites."
#| fig-height: 10
#| fig-width: 10
trajectories %>%
  filter(variable == "ba") %>%
  ggplot(aes(rel_year, y)) +
  geom_line(aes(group = paste(site, plotnum), col = as.character(plotnum))) +
  geom_point(aes(group = paste(site, plotnum), col = as.character(plotnum)),
    data = data %>%
      filter(variable == "ba"), alpha = .5
  ) +
  facet_wrap(~site, scales = "free") +
  theme_bw() +
  xlab("") +
  ylab("") +
  theme(legend.position = "bottom") +
  scale_color_discrete(guide = "none")
```

```{r traj_div}
#| message: false
#| warning: false
#| fig-cap: "Genus diversity jointly inferred trajectories for all sites."
#| fig-height: 10
#| fig-width: 10
sim_sqrt_trans <- function() {
  trans_new(
    "sim_sqrt",
    function(x) sign(x) * sqrt(abs(x)),
    function(x) sign(x) * x^2
  )
}
delta <- filter(parameters, variable == "delta") %>%
  select(site, plot, median) %>%
  rename(delta = median)
continent <- read_tsv("../modelling/outputs/parameters.tsv") %>%
  select(site, region) %>%
  unique() %>%
  na.omit() %>%
  mutate(continent = paste0(str_sub(region, 1, 2), "."))
data_equ <- read_tsv("../modelling/data/derived_data/data_equ.tsv")
g <- trajectories %>%
  filter(variable == "diversity_q1_gen") %>%
  left_join(delta) %>%
  left_join(continent) %>%
  ggplot(aes(rel_year, y, col = delta * 100)) +
  geom_line(aes(group = paste(site, plotnum))) +
  geom_point(aes(group = paste(site, plotnum)),
    data = data %>%
      filter(variable == "diversity_q1_gen") %>%
      left_join(delta) %>%
      left_join(continent),
    alpha = .5
  ) +
  geom_boxplot(
    data = data_equ %>%
      filter(treatment == "control") %>%
      filter(site %in% data$site) %>%
      mutate(rel_year = 61) %>%
      filter(variable == "diversity_q1_gen") %>%
      left_join(delta) %>%
      left_join(continent),
    width = 4, fill = NA, col = "black"
  ) +
  geom_boxplot(
    data = data_equ %>%
      filter(treatment == "logged") %>%
      filter(site %in% data$site) %>%
      mutate(rel_year = -1) %>%
      filter(variable == "diversity_q1_gen") %>%
      left_join(delta) %>%
      left_join(continent),
    width = 4, fill = NA, col = "black"
  ) +
  facet_wrap(~ paste(continent, site), scales = "free_y", nrow = 6) +
  theme_bw() +
  xlab("") +
  ylab("") +
  scale_color_gradient2(expression(delta[~Diversity] ~ "[ % ]"),
    high = muted("red"), mid = "grey", low = muted("blue"),
    transform = "sim_sqrt"
  )
ggsave("figures/fig3.png", width = 10, height = 10, g, dpi = 300, bg = "white")
g
```

## IDH - H1

```{r}
#| message: false
#| warning: false
t <- filter(parameters, variable == "delta") %>%
  left_join(continent) %>%
  mutate(significant = as.numeric(sign(q5) == sign(95))) %>%
  mutate(sign = as.numeric(median > 0)) %>%
  group_by(region, site, sign) %>%
  mutate(median = median * 100) %>%
  summarise(
    N = n(), significant = sum(significant) / N * 100, min = min(median),
    max = max(median)
  ) %>%
  group_by(region, site, sign) %>%
  transmute(cell = paste0(
    N, " - ", round(significant), "% sign. [",
    round(min), ", ", round(max), "]"
  )) %>%
  mutate(sign = ifelse(sign == 0, "negative", "positive")) %>%
  pivot_wider(names_from = sign, values_from = cell)
write_tsv(t, "figures/tab2.tsv")
knitr::kable(t)
```

```{r delta_phi}
#| message: false
#| warning: false
#| fig-cap: "Diveristy hump delta with basal area disturbance intensity phi for each plot in each site." #nolint
filter(parameters, variable == "delta") %>%
  mutate(phiba = filter(parameters, variable == "phiba")$median) %>%
  ggplot(aes(median * 100, (1 - phiba) * 100)) +
  geom_vline(xintercept = 0) +
  geom_segment(aes(
    x = (median - sd) * 100,
    xend = (median + sd) * 100
  ), col = "darkgrey") +
  geom_segment(aes(x = q5 * 100, xend = q95 * 100),
    size = .1, col = "darkgrey"
  ) +
  geom_point() +
  facet_wrap(~site, scales = "free") +
  theme_bw() +
  ylab(expression(phi[BA] ~ "[ % lost ]")) +
  xlab(expression(delta[Diversity] ~ "[ % ]"))
```

```{r delta_site}
#| message: false
#| warning: false
#| fig-cap: "Diveristy hump delta distribution per site."
filter(parameters, variable == "delta") %>%
  ggplot(aes(median)) +
  geom_histogram(fill = "darkgrey", col = NA) +
  geom_vline(xintercept = 0) +
  theme_bw() +
  facet_wrap(~site, scales = "free_y") +
  xlab(expression(delta[Diversity] ~ "[ % ]"))
```

```{r delta_all}
#| message: false
#| warning: false
#| fig-cap: "Diveristy hump delta distribution across sites."
filter(parameters, variable == "delta") %>%
  ggplot(aes(median)) +
  geom_histogram(aes(fill = site), col = "black") +
  geom_vline(xintercept = 0) +
  theme_bw() +
  xlab(expression(delta[Diversity] ~ "[ % ]"))
```

## IDH - H2

```{r gamma}
#| message: false
#| warning: false
#| fig-cap: "Posteriors of the parameter gamma representing basal are disturbance intensity effect on genus diversity short term increase with plot areas in label." #nolint
g <- filter(parameters, variable == "gamma") %>%
  mutate(median = -median, q5 = -q5, q95 = -q95) %>%
  mutate(marked = sign(median - sd) == sign(median + sd)) %>%
  left_join(continent) %>%
  mutate(name = paste(continent, site)) %>%
  ggplot(aes(median, reorder(name, -median), color = median)) +
  geom_vline(xintercept = 0, col = "darkgrey") +
  geom_segment(aes(
    x = (median - sd),
    xend = (median + sd)
  )) +
  geom_segment(aes(x = q5, xend = q95), size = .1) +
  geom_point(aes(size = marked)) +
  theme_bw() +
  xlab(expression(gamma ~ "[" ~ "%"[~Diversity] ~ "%"[~Basal ~ Area] ~ ""^{
    -1
  } ~ "]")) +
  ylab("") +
  scale_color_gradient2(
    guide = "none",
    high = muted("red"), mid = "grey", low = muted("blue"),
    transform = "sim_sqrt"
  ) +
  scale_size_discrete(guide = "none", range = c(1.5, 4))
ggsave("figures/fig4.png", width = 6, height = 4, g, dpi = 300, bg = "white")
g
```

```{r phi_delta_gamma}
#| message: false
#| warning: false
#| fig-cap: "Inferred relationship between disturbance intensity and diversity short term increase." #nolint
pars_s <- filter(parameters, variable == "gamma") %>%
  mutate(delta0 = filter(parameters, variable == "delta0")$median)
filter(parameters, variable %in% c("delta", "phiba")) %>%
  select(site, plot, variable, median) %>%
  pivot_wider(names_from = variable, values_from = median) %>%
  ggplot(aes((1 - phiba), delta)) +
  facet_wrap(~site) +
  geom_abline(
    aes(
      slope = -mean,
      intercept = mean + delta0, group = site
    ),
    data = pars_s
  ) +
  geom_abline(
    aes(
      slope = -mean - sd,
      intercept = mean - sd + delta0, group = site
    ),
    data = pars_s, linetype = "dashed"
  ) +
  geom_abline(
    aes(
      slope = -mean + sd,
      intercept = mean + sd + delta0, group = site
    ),
    data = pars_s, linetype = "dashed"
  ) +
  geom_point() +
  theme_bw() +
  xlab(expression(phi[BA] ~ "[ % lost ]")) +
  ylab(expression(delta[Diversity] ~ "[ % ]")) +
  theme(panel.spacing = unit(0, "lines")) +
  scale_x_continuous(labels = ~ . * 100) +
  scale_y_continuous(labels = ~ . * 100)
```

```{r gamma_all}
#| message: false
#| warning: false
#| fig-cap: "Posteriors of the parameter gamma representing basal are disturbance intensity effect on genus diversity short term increas across sites." #nolint
filter(parameters, variable == "gamma") %>%
  ggplot(aes(median)) +
  geom_histogram(fill = "lightgrey", col = "black", binwidth = 0.2) +
  geom_vline(xintercept = 0) +
  theme_bw() +
  scale_x_reverse(expression(gamma ~ "[" ~ "%" ~ "%"^{
    -1
  } ~ "]"))
```

## Sensitivity to area

```{r}
#| message: false
#| warning: false
g <- read_tsv("../sensitivity/data/derived_data/gamma_rec.tsv") %>%
  select(area, site, variable, median, sd, q5, q95) %>%
  na.omit() %>%
  mutate(median = -median, q5 = -q5, q95 = -q95) %>%
  mutate(marked = sign(median - sd) == sign(median + sd)) %>%
  left_join(continent) %>%
  mutate(name = paste(continent, site)) %>%
  ggplot(aes(paste0(area, "-ha"), median, col = name)) +
  geom_hline(yintercept = 0, col = "darkgrey") +
  geom_segment(aes(y = median - sd, yend = median + sd),
    position = position_nudge(x = c(-0.1, 0.1))
  ) +
  geom_segment(aes(y = q5, yend = q95),
    size = .3,
    position = position_nudge(x = c(-0.1, 0.1))
  ) +
  geom_point(aes(size = marked), position = position_nudge(x = c(-0.1, 0.1))) +
  theme_bw() +
  xlab("") +
  ylab(expression(gamma ~ "[" ~ "%"[~Diversity] ~ "%"[~Basal ~ Area] ~ ""^{
    -1
  } ~ "]")) +
  scale_color_discrete("") +
  theme(
    legend.position = c(0.2, 0.8),
    legend.background = element_rect(fill = NA)
  ) +
  scale_size_discrete(guide = "none", range = c(1.5, 4))
ggsave("figures/fig5.png", width = 6, height = 4, g, dpi = 300, bg = "white")
g
```
