# Sensitivity {.unnumbered}

```{r set}
#| message: false
#| warning: false
library(tidyverse)
library(entropart)
library(cmdstanr)
```

We explored the role of plot area in model sensitivity, particularly with regard to the site parameters representing the linear effect of forest structure disturbance on intermediate diversity extremes. For this, we focused on the Paracou and Mbaiki sites, whose original plots are 6.25 and 4 ha respectively, and virtually subdivided them into adjacent plots of 4, 2, 1, 0.5 and 0.25 ha (for reference, the main analyses were conducted using 1-ha virtual plots). **To update:** We then ran the joint model with basal area and genus diversity using Hill's number of order 1 for each size, and assessed the sign, significance, strength, and uncertainty of the site parameters.

## Data

### Paracou

Paracou plots are 6.25-ha 250x250m plots, we subdivided them into 4-ha, 2-ha, 1-ha, 0.5-ha and 0.25-ha.

```{r subplots_paracou}
#| eval: false
paracou <- "data/raw_data/data_cleaned_paracou.csv" |>
  read_csv() |>
  rename_all(tolower) |>
  rename(x = xtreeplot, y = ytreeplot) |>
  # 6.25-ha
  mutate(plot_625 = plot) |>
  select(site, plot, idtree, x, y, plot_625) |>
  unique() |>
  # 4-ha
  mutate(
    subx = ifelse(x <= 200, x, NA),
    suby = ifelse(y <= 200, y, NA)
  ) |>
  rowwise() |>
  mutate(plot_400 = ifelse(!is.na(subx) && !is.na(suby),
    paste0(plot_625, "_1"), NA
  )) |>
  # 2-ha
  mutate(plot_200 = ifelse(subx > 100, paste0(plot_400, "_2"),
    paste0(plot_400, "_1")
  )) |>
  mutate(subx = ifelse(subx > 100, subx - 100, subx)) |>
  # 1-ha
  mutate(plot_100 = ifelse(suby > 100, paste0(plot_200, "_2"),
    paste0(plot_200, "_1")
  )) |>
  mutate(suby = ifelse(suby > 100, suby - 100, suby)) |>
  # 0.5-ha
  mutate(plot_050 = ifelse(subx > 50, paste0(plot_100, "_2"),
    paste0(plot_100, "_1")
  )) |>
  mutate(subx = ifelse(subx > 50, subx - 50, subx)) |>
  # 0.25-ha
  mutate(plot_025 = ifelse(suby > 50, paste0(plot_050, "_2"),
    paste0(plot_050, "_1")
  )) |>
  mutate(suby = ifelse(suby > 50, suby - 50, suby)) |>
  select(-subx, -suby) |>
  mutate(across(
    c("plot_200", "plot_100", "plot_050", "plot_025"),
    ~ ifelse(is.na(plot_400), NA, .)
  )) |>
  gather(area, subplot, -site, -plot, -idtree, -x, -y) |>
  separate(area, c("x1", "area"), convert = TRUE) |>
  select(-x1) |>
  mutate(area = area / 100)
write_tsv(paracou, "data/derived_data/paracou_subplots.tsv")
```

```{r subplots_paracou_fig}
#| message: false
#| warning: false
#| fig-cap: "Paracou sub-plots divisions."
read_tsv("data/derived_data/paracou_subplots.tsv") |>
  filter(plot == 1) |>
  ggplot(aes(x, y, col = subplot)) +
  geom_point() +
  facet_wrap(~ paste0(area, "-ha")) +
  scale_color_discrete(guide = "none") +
  theme_bw() +
  xlab("") +
  ylab("")
```

### Mbaiki

Mbaiki plots are 4-ha 200x200m plots, we subdivided them into 2-ha, 1-ha, 0.5-ha, and 0.25-ha.

```{r subplots_mbaiki}
#| eval: false
mbaiki <- "data/raw_data/data_cleaned_mbaiki.csv" |>
  read_csv() |>
  rename_all(tolower) |>
  rename(x = xtreeplot, y = ytreeplot) |>
  # 4-ha
  mutate(plot_400 = plot) |>
  select(site, plot, idtree, x, y, plot_400) |>
  unique() |>
  mutate(subx = x, suby = y) |>
  # 2-ha
  mutate(plot_200 = ifelse(subx > 100, paste0(plot_400, "_2"),
    paste0(plot_400, "_1")
  )) |>
  mutate(subx = ifelse(subx > 100, subx - 100, subx)) |>
  # 1-ha
  mutate(plot_100 = ifelse(suby > 100, paste0(plot_200, "_2"),
    paste0(plot_200, "_1")
  )) |>
  mutate(suby = ifelse(suby > 100, suby - 100, suby)) |>
  # 0.5-ha
  mutate(plot_050 = ifelse(subx > 50, paste0(plot_100, "_2"),
    paste0(plot_100, "_1")
  )) |>
  mutate(subx = ifelse(subx > 50, subx - 50, subx)) |>
  # 0.25-ha
  mutate(plot_025 = ifelse(suby > 50, paste0(plot_050, "_2"),
    paste0(plot_050, "_1")
  )) |>
  mutate(suby = ifelse(suby > 50, suby - 50, suby)) |>
  select(-subx, -suby) |>
  gather(area, subplot, -site, -plot, -idtree, -x, -y) |>
  separate(area, c("x1", "area"), convert = TRUE) |>
  select(-x1) |>
  mutate(area = area / 100)
write_tsv(mbaiki, "data/derived_data/mbaiki_subplots.tsv")
```

```{r subplots_mbaiki_fig}
#| message: false
#| warning: false
#| fig-cap: "Mbaiki sub-plots divisions."
read_tsv("data/derived_data/mbaiki_subplots.tsv") |>
  filter(plot == 11) |>
  ggplot(aes(x, y, col = subplot)) +
  geom_point() +
  facet_wrap(~ paste0(area, "-ha")) +
  scale_color_discrete(guide = "none") +
  theme_bw() +
  xlab("") +
  ylab("")
```

### Trajectories

```{r prep_model_data}
#| eval: false
paracou <- read_csv("data/raw_data/data_cleaned_paracou.csv") |>
  rename_all(tolower) |>
  select(site, year, idtree, genus_cleaned, diameter_cor)
mbaiki <- read_csv("data/raw_data/data_cleaned_mbaiki.csv") |>
  rename_all(tolower) |>
  select(site, year, idtree, genus_cleaned, diameter_cor)
raw <- bind_rows(
  read_tsv("data/derived_data/paracou_subplots.tsv"),
  read_tsv("data/derived_data/mbaiki_subplots.tsv")
) |>
  select(area, site, plot, subplot, idtree) |>
  left_join(bind_rows(paracou, mbaiki), relationship = "many-to-many") |>
  filter(!is.na(subplot))
diversity <- raw |>
  subset(!is.na(genus_cleaned)) |>
  group_by(site, area, plot, subplot, year, genus_cleaned) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(site, area, plot, subplot, year) |>
  summarise(
    diversity = Diversity(count,
      q = 0, Level = 100
    )
  )
ba <- raw |>
  group_by(site, area, plot, subplot, year) |>
  summarise(ba = sum(pi * (diameter_cor / 200)^2))
forest <- left_join(diversity, ba) |>
  mutate(rel_year = year - 1986) |>
  group_by(area) |>
  mutate(sitenum = as.numeric(as.factor(site))) |>
  mutate(plot = as.character(plot)) |>
  left_join(read_tsv("data/raw_data/metadata.tsv")) |>
  filter(treatment == "Logging") |>
  group_by(area) |>
  mutate(subplotnum = as.numeric(as.factor(paste(site, subplot))))
write_tsv(forest, "data/derived_data/sensitivity_data.tsv")
data_equ <- forest |>
  group_by(area) |>
  filter(rel_year <= 0) |>
  select(area, site, sitenum, plot, subplot, subplotnum, year, rel_year, diversity, ba)
write_tsv(data_equ, "data/derived_data/sensitivity_data_equ.tsv")
data_rec <- forest |>
  filter(rel_year > 2)
dist_obs <- data_rec |>
  group_by(area, site, subplot) |>
  summarise(ba_post = min(ba, na.rm = TRUE)) |>
  left_join(data_equ |>
    filter(rel_year <= 0) |>
    group_by(area, site, subplot) |>
    summarise(ba_pre = median(ba, na.rm = TRUE))) |>
  mutate(dist_obs = (ba_pre - ba_post) / ba_pre) |>
  select(area, site, subplot, dist_obs) |>
  mutate(dist_obs = ifelse(dist_obs < 0, 0, dist_obs))
data_rec <- data_rec |>
  select(area, site, sitenum, plot, subplot, subplotnum, year, rel_year, diversity, ba) |>
  left_join(dist_obs)
write_tsv(data_rec, "data/derived_data/sensitivity_data_rec.tsv")
```

```{r traj_div}
#| message: false
#| warning: false
#| fig-cap: "Diversity trajectories."
read_tsv("data/derived_data/sensitivity_data.tsv") |>
  ggplot(aes(rel_year, diversity, col = subplot)) +
  geom_line() +
  geom_point() +
  facet_grid(site ~ paste0(area, "-ha"), scales = "free_y") +
  theme_bw() +
  scale_color_discrete(guide = "none") +
  geom_vline(xintercept = 0) +
  xlab("") +
  ylab(expression("Genus diversity" ~ q == 1))
```

## Fits

```{r fit}
#| eval: false
areas <- c(6.25, 4.00, 2.00, 1.00, 0.50, 0.25)
recovery <- cmdstan_model("models/model.stan")
for (a in areas) {
  data_equ <- read_tsv("data/derived_data/sensitivity_data_rec.tsv") |>
    filter(area == a)
  data_rec <- read_tsv("data/derived_data/sensitivity_data_rec.tsv") |>
    filter(area == a)
  ind <- data_rec |>
    select(area, site, plot, subplot, sitenum, subplotnum, dist_obs) |>
    unique()
  mdata <- list(
    n = nrow(data_rec),
    s = max(data_rec$sitenum),
    p = max(data_rec$subplotnum),
    y = data_rec$diversity,
    t = data_rec$rel_year,
    site = data_rec$sitenum,
    plot = data_rec$subplotnum,
    site_plot = ind$sitenum,
    n_u = nrow(data_equ),
    y_u = data_equ$diversity,
    plot_u = data_equ$subplotnum,
    dist_p = ind$dist_obs
  )
  dir <- file.path("chains", paste0("rec", "_", a * 100))
  unlink(dir, recursive = TRUE)
  dir.create(dir)
  fit <- recovery$sample(
    mdata,
    chains = 4,
    parallel_chains = 4,
    save_warmup = FALSE,
    refresh = 100,
    output_dir = dir
  )
}
```

```{r res}
#| eval: false
fits <- tibble(area = c(6.25, 4.00, 2.00, 1.00, 0.50, 0.25)) |>
  mutate(directory = file.path("chains", paste0("rec", "_", area * 100))) |>
  rowwise() |>
  mutate(files = list(list.files(directory, full.names = TRUE))) |>
  mutate(fit = list(as_cmdstan_fit(files)))
time <- fits |>
  mutate(time = list(fit$time()[[2]])) |>
  select(area, time) |>
  unnest(time)
write_tsv(time, "data/derived_data/sensitivity_time_rec.tsv")
gamma <- fits |>
  mutate(gamma = list(fit$summary(c("gamma")))) |>
  select(area, gamma) |>
  unnest(gamma)
write_tsv(gamma, "data/derived_data/sensitivity_gamma_rec.tsv")
data_rec <- read_tsv("data/derived_data/sensitivity_data_rec.tsv")
ind <- data_rec |>
  select(area, site, plot, subplot, sitenum, subplotnum, dist_obs) |>
  unique()
delta <- fits |>
  mutate(delta = list(fit$summary("delta_p"))) |>
  select(area, delta) |>
  unnest(delta) |>
  separate(variable, c("variable", "p", "subplotnum"), convert = TRUE) |>
  select(-p) |>
  left_join(ind)
write_tsv(delta, "data/derived_data/sensitivity_delta_rec.tsv")
```

### Time

```{r time}
#| message: false
#| warning: false
#| fig-cap: "Duration recovery model."
read_tsv("data/derived_data/sensitivity_time_rec.tsv") |>
  ggplot(aes(paste0(area, "-ha"), total / 60)) +
  geom_boxplot() +
  theme_bw() +
  xlab("") +
  ylab("Time [ m ]") +
  scale_y_log10()
```

### $\gamma$

```{r gamma}
#| message: false
#| warning: false
#| fig-cap: "Caption."
read_tsv("data/derived_data/sensitivity_gamma_rec.tsv") |>
  select(area, variable, median, sd, q5, q95) |>
  na.omit() |>
  ggplot(aes(paste0(area, "-ha"), median)) +
  geom_hline(yintercept = 0, col = "darkgrey") +
  geom_segment(aes(y = median - sd, yend = median + sd),
    position = position_nudge(x = c(-0.1, 0.1))
  ) +
  geom_segment(aes(y = q5, yend = q95),
    size = .3,
    position = position_nudge(x = c(-0.1, 0.1))
  ) +
  geom_point(position = position_nudge(x = c(-0.1, 0.1))) +
  theme_bw() +
  xlab("") +
  ylab(expression(gamma ~ "[" ~ "%" ~ "%"^{
    -1
  } ~ "]")) +
  scale_color_discrete("")
```

### $\delta$

```{r delta}
#| message: false
#| warning: false
#| fig-cap: "Caption."
read_tsv("data/derived_data/sensitivity_delta_rec.tsv") |>
  na.omit() |>
  ggplot(aes(dist_obs * 100, median * 100)) +
  geom_point(aes(col = site)) +
  geom_smooth(method = "lm", formula = y ~ 1 + x) +
  theme_bw() +
  facet_wrap(~ paste0(area, "-ha")) +
  xlab(expression(dist[BA] ~ "observed [ % lost ]")) +
  ylab(delta[~Diversity] ~ "[ % ]") +
  ggpubr::stat_cor()
```
